Brief Article

### 2017-05-31 latex/secretkeybasedpublickeyencryption/paper.tex
uploading new design

\author{Ella Rose}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[An Efficient Public Key Encryption Algorithm]{An Efficient Public Key Encryption Algorithm}

@@

% use optional argument because the \LaTeX command breaks the PDF keywords
\keywords[Homomorphic Encryption, Symmetric Encryption, Asymmetric Encryption]{Homomorphic Encryption, Symmetric Encryption, Asymmetric Encryption}

\begin{abstract}
 We define a homomorphic encryption based public-key encryption construction that is efficient in terms of key size, ciphertext size, key generation time, and encryption/decryption time. The construction requires only addition, multiplication, and modular reduction on large integers. The security of the secret key cipher that underlies the scheme appears to rely only on the distribution of the random numbers generated and the correct selection of key parameters.
 \end{abstract}

@@

\section{Introduction}
 Ciphers that support homomorphic operations on ciphertexts are very attractive because they tend to make the creation of public key encryption schemes relatively straightforward and simple. This paper presents such a cipher, as well as a public key encryption system based on it.

\section{Definitions}
We use $+$ to denote addition, $a * b$ to denote multiplication, $a / b$ to denote division, and $\mod p$ to denote the modulo operation.

@@

A fully homomorphic cipher supports multiple operations on ciphertexts, usually addition and multiplication. The public key encryption construction presented in this paper does not require a fully homomorphic cipher, only a a partially homomorphic cipher. We mention it here only for the sake of completeness.

@@

\subsection{Key Generation}
Generation of a public key is very simple: Distribute one encryption of $1$, as well an encryption of $0$. The size of the public key will be determined by the size of the ciphertexts generated by the secret key cipher.

\begin{flalign*}
p_1 := encrypt(1, p_k) \\
p_2 := encrypt(0, p_k) \\
key_{public} = (p_1, p_2)
\end{flalign*}

\subsection{Encryption}
To encrypt a value: First the encryption of $1$ is multiplied by the plaintext message; Then, the encryption of $0$ is multiplied by a random plaintext integer; Finally, these products are summed modulo n to produce the ciphertext.

\begin{flalign*}
ciphertext := (p_1 * m) + (p_2 * r) \mod n
\end{flalign*}

\subsection{Decryption}
Decryption simply applies the decryption circuit of the secret key cipher to the plaintext. The result will be equivalent to the message, because:

\begin{flalign*}
D(ciphertext) \equiv (1 * m) + (0 * r_1) \equiv m
\end{flalign*}

\section{Secret Key Cipher}
\subsection{Basic Concept}
Points on a lattice can be added together, and this effect is preserved even when a modular reduction step is applied, as long as the lattice basis and modulus are coprime. We exploit this modular reduction to both provide security and decrease the size of ciphertexts considerably.

\subsubsection{Storing Data In the Point Or In the Error}
When using a lattice to secretly store information, there are two places that the information can be stored: Either the information is encoded as a point on the lattice, or the information is the distance to the nearest lattice point.\\

Written in notation, a point on a lattice takes the form $p * q$, while a point plus some error takes the form $p * q + e$. In the former, the information is recovered by dividing by $p$ - or alternatively, as in this case, multiplying by the modular multiplicative inverse of $p \mod n$. In the latter, the information is recovered by dividing by $p$ and taking the remainder. \\

We need to be able to encrypt 0, so using $p * q$ where $q = m$ is not suitable - therefore we must use $p * q + e$ where $e = m$. By itself, this cipher would be insecure; However, it does output a randomly distributed value. We then encrypt this randomly distributed value as a point on the lattice to produce the ciphertext.

\subsection{Encryption}
\begin{flalign*}
r_2 := (k_1 * r_1) + m\\
c := k_2 * r_2 \mod n
\end{flalign*}

\subsection{Decryption}
\begin{flalign*}
k_i := inverse(k_2, n)\\
r_2 := c * k_i \mod n\\
m := r_2 \mod k_1
\end{flalign*}

\subsection{Key Generation}
$k_1$ must be selected such that it is greater then the largest message that will be encrypted. If 256-bit messages are to be encrypted, then $k_1$ must be greater then 256 bits in size. We elect to utilize 264 bits for our $k_1$ values. \\

If the modulus is significantly larger then $k_2$, then $k_2$ could be determined by taking the gcd of $k_2 r_i$ pairs that happened to not reduce. This can be remedied by ensuring that $k_2$ is large enough. We use choose $k_2$ by generating a 256-bit number, and subtracting it from the modulus to obtain $k_2$. Selecting the point starting from $n$ instead of $0$ ensures that $k_2$ is large enough. If $k_2 >= n - k_2$, then gcd cannot be used to extract the value of $k_2$.

\section{Security Analysis}
\subsection{Secret Key Cipher}
The security of the secret key cipher seems to reduce almost exclusively to the distribution of values produced by $(k_1 * r_1) + m$. If $(k_1 * r_1) + m$ is considered to be a uniformly random value, then the secret key cipher is effectively equivalent to a One Time Pad, with a re-usable key. This seems implausible, and so we consider that $(k_1 * r_1) + m$  probably does not always follow such a uniformly random distribution in all cases.

@@

\subsection{Public Key Cipher}
The security of the public key system seems very straightforward: as long as $p_2 * r_1 \mod n$ produces a random value, then $(p_1 * m) + (p_2 * r_1) \mod n$ is effectively equivalent to applying a random key to an encoding of the message. As long as the key material is unpredictable and remains unknown, then there shold be no way to recover the message, given only a ciphertext and the public key.

@@

With our public domain python implementation, we can encrypt 10,000 256-bit messages in approximately 0.58 seconds, while we can decrypt 10,000 ciphertexts in approximately 0.32 seconds. While a python implementation may not be the fastest possible, it performs well as much of the computation time is spent on BigNum operations anyways. \\

The private key size could be decreased a small amount at the expense of computation time; The figure presented in the paragraph prior represents the storage of $k_1, k_2, inverse(k_2, n)$. It is not necessary to store both $k_2$ and $inverse(k_2, n)$, as they can be calculated from each other. Considering the small amount of space it costs to store the inverse versus how long it takes to compute it, we choose to cache this value.

@@

Additional code is required for supporting functions such as the generation of prime numbers (only if $n$ is being generated/not already established), computation of modular inverses, and unit testing. \\

Assuming that BigNum operations can be implemented in a manner that does not leak timing information, it appears that the construction is resistant to timing side channel analysis.

\section{Resource Asymmetry}
There is a computational asymmetry between encryption and decryption, even in the simplest setting. Encryption takes longer then decryption, and this is something that potentially could be tuned or exploited for other cryptographic constructions. Perhaps there is a way to distribute large public keys, which could be used as a part of a proof-of-space or proof-of-work scheme, or perhaps password hashing.

>>>>

\author{Ella Rose}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[A Practical Public Key Encryption Algorithm]{A Practical Public Key Encryption Algorithm}

@@

% use optional argument because the \LaTeX command breaks the PDF keywords
\keywords[Homomorphic Encryption, Public Keyc Encryption]{Homomorphic Encryption, Public Key Encryption}

\begin{abstract}
 We define a public-key encryption construction that is efficient in terms of key size, ciphertext size, key generation time, and encryption/decryption time. The construction requires only addition, multiplication, and modular reduction on large integers. The security of the scheme appears to rely only on the distribution of the random numbers generated and the correct selection of key parameters.
 \end{abstract}

@@

\section{Introduction}
 Ciphers that support homomorphic operations on ciphertexts are very attractive because they tend to make the creation of public key encryption schemes relatively straightforward and simple. This paper presents such a public key encryption system, as well as a secret key cipher that is suitable for instantiating the system.

\section{Definitions}
We use $+$ to denote addition, $a * b$ to denote multiplication, $a / b$ to denote division, and $\mod p$ to denote the modulo operation. We use the function $inverse(k, n)$ to indicate the computation of the modular multiplicative inverse of $k \mod n$.

@@

A fully homomorphic cipher supports multiple operations on ciphertexts, usually addition and multiplication. The public key encryption construction presented in this paper does not require a fully homomorphic cipher, only a a partially homomorphic cipher. We mention the term here only for the sake of completeness.

@@

\subsubsection{Key Generation}
Generation of a public key is very simple: Distribute one encryption of $1$, as well an encryption of $0$. The size of the public key will be determined by the size of the ciphertexts generated by the secret key cipher.

\begin{flalign*}
p_0 := encrypt(0, p_k) \\
p_1 := encrypt(1, p_k) \\
key_{public} = (p_0, p_1)
\end{flalign*}

\subsubsection{Encryption}
To encrypt a value: First the encryption of $1$ is multiplied by the plaintext message; Then, the encryption of $0$ is multiplied by a random plaintext integer; Finally, these products are summed modulo n to produce the ciphertext.

\begin{flalign*}
ciphertext := (p_0 * r) + (p_1 * m) \mod n
\end{flalign*}

\subsubsection{Decryption}
Decryption simply applies the decryption circuit of the secret key cipher to the plaintext. The result will be equivalent to the message, because:

\begin{flalign*}
D(ciphertext) \equiv (0 * r) + (1 * m)\equiv m
\end{flalign*}

\subsection{Secret Key Cipher}
\subsubsection{Basic Concept}
Points on (and off of) a lattice can be added together, and this effect is preserved even when a modular reduction step is applied, as long as the lattice basis and modulus are coprime and the point is smaller then the modulus. We exploit this modular reduction to both provide security and decrease the size of ciphertexts considerably.

\subsubsection{Storing Data In the Point Or In the Error}
When using a lattice to secretly store information, there are two places that the information can be stored: Either the information can be encoded as a point on the lattice, or the information can be encoded as the distance to the nearest lattice point.\\

Written in notation, a point on a lattice takes the form $p * q$, while a point plus some error takes the form $p * q + e$. In the former, the information is recovered by dividing by $p$ - or alternatively, multiplying by the modular multiplicative inverse of $p \mod n$. In the latter, the information is recovered by dividing by $p$ and taking the remainder. \\

We need to be able to encrypt 0, and using $p * q$ where $q = m$ is not suitable - therefore we must use $p * q + e$ where $e = m$. By itself, this cipher would be insecure; However, it does output a randomly distributed value. We use this randomly distributed value as a point on a secret lattice to produce the ciphertext.

\subsubsection{Encryption}
\begin{flalign*}
r_2 := (k_1 * r_1) + m\\
c := k_2 * r_2 \mod n
\end{flalign*}

\subsubsection{Decryption}
\begin{flalign*}
k_i := inverse(k_2, n)\\
r_2 := c * k_i \mod n\\
m := r_2 \mod k_1
\end{flalign*}

\subsubsection{Key Generation}
$k_1$ must be selected such that it is greater then the largest message that will be encrypted. If 256-bit messages are to be encrypted, then $k_1$ must be greater then 256 bits in size. We elect to utilize 264 bits for our $k_1$ values. \\

If the modulus is significantly larger then $k_2$, then $k_2$ could be determined by taking the gcd of $k_2 r_i$ pairs that happened to not reduce. This can be remedied by ensuring that $k_2$ is large enough. We use choose $k_2$ by generating a 256-bit number, and subtracting it from the modulus to obtain $k_2$. Selecting the point starting from $n$ instead of $0$ ensures that $k_2$ is large enough. If $k_2 >= n - k_2$, then gcd cannot be used to extract the value of $k_2$ even for consecutive small $q$.

\section{Security Analysis}
\subsection{Secret Key Cipher}
The security of $k * q \mod n$ as a cipher seems to reduce almost exclusively to the distribution of values for $q$, which is determined by $(k_1 * r_1) + m$. If $(k_1 * r_1) + m$ is considered to be a uniformly random value, then the secret key cipher is effectively equivalent to a One Time Pad, with a re-usable key. This is demonstrably not the case, as perfect secrecy is not provided by this cipher - A given ciphertext decrypts to one and only one particular plaintext message. We consider that $(k_1 * r_1) + m \mod n$  may not always follow a uniformly random distribution.

@@

Encryptions of 0 take the form $p * k * q \mod n$. Given many such samples for random $q$, can p be recovered? If $x := p * k$, then the expressions look like $x * q \mod n$ for random $q$, which should be secure on the assumption that $q$ is unpredictable. If $q$ is unpredictable, then it should not be possible to recover $p$ from encryptions of $0$.\\

Encryptions of $m$ take the form $(p * k * q) + p * m \mod n$. Given many such samples for random q, can $p$ be recovered? If $p * k * q \equiv p * y \mod n$, and $y \mod n$ is unpredictable, then it should not be possible to recover $p$ from encryptions of $m$.\\

Under chosen ciphertext attack, we can obtain the value of $p_i \mod k$ by submitting the value of $1$ as a ciphertext to the decryption operation. Similarly, we can obtain the value of $p_i * c \mod n \mod k$ for some chosen values $c$. It is not yet clear how this information can be leveraged to recover $p$ or $p_i$.\\

If $k$ is a power of two, then $p_i \mod k$ directly reveals some of the lower bits of $p_i$. It is fantastically unlikely that $k$ will be a power of two if it is a randomly generated 264-bit number.

\subsection{Public Key Cipher}
The security of the public key system seems very straightforward: as long as $p_0 * r \mod n$ produces an unpredictable random value, then $(p_1 * m) + (p_0 * r) \mod n$ is effectively equivalent to applying a random key to an encoding of the message. As long as the key material is unpredictable and remains unknown, then, given only a ciphertext and the public key, there shold be no way to recover the plaintext message.

@@

With our public domain python implementation, we can encrypt 10,000 256-bit messages in approximately 0.58 seconds, while we can decrypt 10,000 ciphertexts in approximately 0.32 seconds. While a python implementation may not be the fastest possible, it performs well as much of the computation time is spent on BigNum operations anyways.\\

The public key encryption operation requires effectively the same amount of time as the secret key encryption operation; Consequently, there is little motivation to use the secret cipher as an encryption algorithm by itself. \\

The private key size could be decreased a small amount at the expense of computation time; The figure cited above represents the storage of $k_1, k_2, inverse(k_2, n)$. It is not necessary to store both $k_2$ and $inverse(k_2, n)$, as they can be calculated from each other. Considering the small amount of space it costs to store the inverse versus how long it takes to compute it, we choose to cache this value.

@@

Additional code is required for supporting functions such as the generation of prime numbers (only if $n$ is being generated/not already established), computation of modular inverses, and unit testing. These are modest requirements that are already well facilitated by existing cryptographic libraries, and in other cases can be implemented without much complexity. \\

Assuming that BigNum operations can be implemented in a manner that does not leak timing information, it appears that a naive implementation of the algorithm is resistant to timing side channel analysis, as there are no tables or branches.

\subsection{Resource Asymmetry}
There is a computational asymmetry between encryption and decryption: Encryption takes longer then decryption. This is something that potentially could be tuned or exploited for other cryptographic constructions. Perhaps there is a way to distribute large public keys, which could be used as a part of a proof-of-space or proof-of-work scheme, or perhaps password hashing.

### 2017-05-31 latex/secretkeybasedpublickeyencryption/scratchtest.tex
uploading new design

%%% END Article customizations

@@

\begin{document}
\maketitle

\section{Cipher Constructions}

The design of a cipher can be split into (at least) two parts: The cipher construction, and the core permutation. The two archetypal examples of cipher construction are the Feistel Network and Substiution-Permutation Network.

In the former, the state is divided into two sections that are typically of equal size; A key is added to some side(s), then a core "randomizing" function is applied to one half to produce a key stream; This key stream is then combined with the other half, in order to encipher it. We can see then, that the core "randomizing function" is interchangeable within the Feistel Network construction.

In a Substitution-Permutation network, the state is typically subdivided into many smaller sections. The algorithm typically consists of addition of a key layer, followed by at least one application of an invertible "randomizing" function applied to the entire state; This is repeated until a sufficient "security margin" is obtained. This construction can be modeled as the composition of two functions: The key addition function, and the application of the core permutation on the state. As in the Feistel Network, the core permutation is interchangeable within the Substitution-Permutation network construction.

We can see then, that there are at least two, effectively separate aspects to cipher design: The cipher construction, and the core permutation. This work focuses on the former. While the core permutation is critical for resistance to generic attacks such as linear and differential cryptanalysis, there also exist generic attacks against the cipher constructions. Additionally, while the design of the permutation can be hugely influential on the performance of the algorithm, so too can the cipher construction.

\section{Construction Design Considerations}

First, the target platform needs to be established. What is ideal for constrained hardware is not necessarily ideal for a modern CPU. In addition to the target platform(s), the performance goals of the algorithm should be outlined: Is latency important, or is throughput for large messages important? Is it important to use as little (or much) RAM?

Less obvious concerns that are still very important include: Is resistance to side channel leakage minimized? Does it facilitate or impede attempts to implement counter-measures?

"The simple" - RC4 like (shuffling huge array of words)

"The small" - 8-bit friendly/constrained device friendly.

"The fast" - salsa like (SIMD with big registers)

\subsection{A subsection}

More text.

>>>>

\usepackage{mathtools}
%%% END Article customizations

@@

\begin{document}
\begin{flalign*}
P := big\_prime\\
p_0 := k_2 (k_1 r_1 + 0) \mod P\\
p_1 := k_2 (k_1  r_2 + 1) \mod P\\
key_{public} := (p_0, p_1)\\
key_{private} := (k_1, k_2, inverse(k_2, P))\\
\\
c := p_0 r + p_1 m \mod P\\
m := k_{2i} c \mod P \mod k_1
\end{flalign*}
